Can government public communications elicit undue trust? Exploring the interaction between symbols and substantive information in communications

### 2019-03-08 paper2_JPART_rev_01.Rmd
Add files via upload

---
title: "Can government public communications elicit undue trust? Exploring the interaction between symbols and substantive information in communications"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
urlcolor: blue
---

@@

```{r , include=FALSE, echo=FALSE}
#load("SVIVA_R_ENV.RData")
source("SVIVA_exp2_dm_03.R")

@@

Public administration scholars are highly concerned about the implications of citizens’ low trust in government, and their underestimation of public-service performance. One of the main remedies proposed for this is greater investment in direct government communications to citizens. Congruently, over the past generation, economically-developed democratic countries have witnessed an increase in government public communications, specifically through advertising, social media, and e-government platforms [@eshuis_2012; @marland_2017; @mergel_2013; @mickoleit_2014]. In the US, for example, federal government agencies spend approximately 1.5 Billion dollars annually on contracts related to public relations and advertising activities in traditional and digital media [@unitedstatesgovernmentaccountabilityoffice_2017].

Extant public administration research on the effect of government public communications on citizens attitudes has mostly focused on the *informative* aspects of these communications, whereby government organizations provide citizens with evidence about their performance and details and explanations about their policies and actions [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @porumbescu_2018; @grimmelikhuijsen_2015; @ho_2017]. Public communications are expected to increase citizens' trust insofar as they provide citizens with favorable information on the performance of organizations and logically persuasive explanations for their policies. At the same time, experiment-based studies that explored the responses of citizens to substantive information revealed cognitive biases that distort their processing of information in communications, and undermine its effectiveness [@baekgaard_2016; @grimmelikhuijsen_2013; @grimmelikhuijsen_2014; @hvidman_2016; @james_2011; @james_2014; @james_2017; @marvel_2015; @olsen_2017; @james_2017a; @grimmelikhuijsen_2017].

@@

The theoretical hypotheses are graphically displayed in **Figure 1**.

<br>
**Figure 1: Theoretical model**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/theoretical_model.png) 
<br>

@@

Methodologically, the EPM is a good candidate for this study due to its significant investment, over the past years, in public communications. In these communications, the EPM used an array of symbolic elements, which became widely recognized by the public. Here, I focused on three prominent symbolic elements used in the EPM's public communications, and displayed below in **Figure 2**. The first symbolic element is the unique brand logo of the EPM (Figure 2a). This logo consists of a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second symbolic element which I focused on, is the color green, which is strongly associated with environmental protection. The EPM logo and the green color are used in almost all of EPM’s visual communications, and they represent key elements of its unique "visual identity" [@vandenbosch_2005]. The third element is celebrity endorsements. Specifically, I selected two high-profile celebrity comedians, *Tal Friedman* and *Ido Rosenblum*, who previously endorsed two salient advertising campaigns of the EPM. The campaign starring the comedian Tal Friedman (Figure 2b) was launched in 2010, and included a series of advertisements in which Friedman was presented with a head covered in green grass, and accordingly the campaign's slogan was: "starting to think green". The campaign starring the comedian Ido Rosenblum (Figure 2c) was launched in 2017, and focused on reducing the use of disposable bags.[^footnote_campaigns] The two campaigns were highly successful in terms of their recognizability and likability, as suggested by their evaluation reports by the Government advertising agency.

@@

<br>
**Figure 2: Symbolic elements of EPM**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/EPM_symbols.png) 
<font size="0.5">

@@

Altogether, for each policy plan, the survey experiment has a factorial design of 3 (symbols: real symbols, fake-symbols, no-symbols) x 2 (information: strong vs. weak policy). The assignment to the different experimental groups is graphically presented in **Figure 3**. The experimental groups are balanced with respect to demographics, and pre-manipulation variables. Throughout the rest of the paper, I will analyze the results at the policy level, in order to simplify the analyses. Accordingly, I transformed the dataset into a multilevel structure of policies nested within subjects.[^footnote_combined_analyses]

@@

<br>
**Figure 3: Experimental design for policy plan communications**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/experiment_flow.png)

@@

In Model 1.2, the effect of the symbols remains intact, and the effect of the weak policy condition is negative and significant (*Coef* = `r mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]). In Model 1.3, the interaction between weak policy and real symbols is positive and significant, whereas the interaction with the fake symbols is insignificant. The interpretation of this model suggests that the real symbols had a greater effect on trust when the policy plan was weak (*Coef* = `r mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`] and `r mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`] compared with the no symbols and fake symbols, respectively). When added to the strong policy plan, the effect of the real symbols is weaker, and significant only compared with the no symbols condition (*Coef* = `r mod_1.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)`, *p*<.1, one-tailed). Congruently, the symbols decreased the differences in trust between the strong and weak policies by `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())/(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())*100)%>%round(0)`%, compared with the no symbols group (from `r mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`] to `r mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`]). The interaction between symbols and information significantly contributes to the model fit (Chi2 = `r anova(mod_1.3,mod_1.2)%>%tidy()%>%filter(term=="mod_1.3")%>%select(statistic)%>%max()%>%round(2)`, *p*<.05). To facilitate the interpretation of this interaction, **Figure 4** provides a descriptive demonstration of the effects of the real symbols on trust in policy, under weak and strong information quality.

**Figure 4: The effect of symbolic elements on trust under weak vs. strong information**

@@

The interaction between symbols and information can also be tested *within-subjects*, by comparing the averaged delta across the different symbols conditions. Based on my second hypothesis, I expect that the averaged delta will be lower among subjects in the real symbols condition. This comparison is graphically displayed in **Figure 5**. The averaged delta in the real symbols is significantly lower than in the no symbols and fake symbols control groups (differences are `r (x_realsymbol_delta-x_nosymbol_delta)%>%abs()%>%round(2)` [`r t_delta_symbol_nosymbol[["conf.int"]][1]%>%round(2)`,`r t_delta_symbol_nosymbol[["conf.int"]][2]%>%round(2)`] and `r (x_realsymbol_delta-x_fakesymbol_delta)%>%abs()%>%round(2)` [`r t_delta_symbol_fakesymbol[["conf.int"]][1]%>%round(2)`,`r t_delta_symbol_fakesymbol[["conf.int"]][2]%>%round(2)`], *t*= `r t_delta_symbol_nosymbol[["statistic"]][["t"]]%>%round(3)` and `r t_delta_symbol_fakesymbol[["statistic"]][["t"]]%>%round(3)`, respectively). The differences between the two control groups are insignificant. These results further support my findings above that familiar symbols decrease subjects' inclination to differentiate between the strong and weak policy plans.

**Figure 5: The effect of symbolic elements on within-subject delta between strong and weak policies**

@@

In the following section, I test the postulation that the interaction between symbols and information is explained or mediated by a decrease in elaboration. To reiterate, in the survey, I used reaction time (log) and memory score as measures for participants' level of elaboration of the policy plans. Based on my third hypothesis, I expect to find a negative effect of the real symbols on these two measures. Thereafter, my fourth hypothesis entails moderated mediation relations: the decrease in elaboration mediates the effect of symbols on trust under the weak policy, but not under the strong policy (see Figure 1).

@@

In all four models, the coefficients of the real symbols are negative and significant, and the coefficients of the fake symbols are insignificant. The real symbols decreased the number of second subjects spent on the pages of the policy plans in the online survey, compared with both control groups.[^footnote_RT_symbols] The effect is significant at the 95% level compared with the no symbols condition (*Coef* = `r mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`]), yet it is not sufficiently significant compared with the fake symbols. The averaged memory score of subjects who saw the real symbols was significantly lower, compared with both the no logo and the fake logo groups (*Coef* = `r mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`], and `r mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`], respectively).[^footnote_memory_symbols] There is no significant change to the results when adding the effects of the information manipulation (in Models 2.2, 2.4). The coefficients of the weak policy in these models suggest that those who saw the fictitious, weak policy plans, spent more time on their pages, yet were less likely to state items that were included in them. **Figure 6** displays the descriptive comparison of the elaboration measures across the symbols conditions. The left side shows its effect on reaction time (logged), and the right side shows its effect on the memory-test scores.

@@

**Figure 6: The effect of symbolic elements on elaboration**

@@

The results of the within-subjects mediation analysis similarly supports the moderated mediation hypothesis. The negative effect of the real symbols on the delta within subjects (between trust in the strong and weak policies) is significantly mediated by the negative effect of the real symbols on reaction time and the memory score. With respect to the reaction time, the estimated proportional mediation is `r (table_med_RT__nosymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% compared with the no symbols, and `r (table_med_RT__fakesymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% compared with the fake symbols. Regarding the memory score, the estimated proportional mediation is `r (table_med_memory__nosymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% and `r (table_med_memory__fakesymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`%, respectively.[^footnote_mediation_sensitivity]

@@

Keeping these limitations in mind, this study has several important implications for public administration research. Most importantly, my findings highlight the significant role of branding and symbolic communication in shaping citizens' attitudes about government. Citizens' responses to symbolic elements in communications thus far received relatively little attention in Public administration literature, compared with the abundant research on their responses to information. Yet, this study, and others, demonstrated that symbols can be fairly effective in increasing citizens' trust, and mitigating their negative attitudes [@alonbarkat_2017; @karens_2016; @teodoro_2018]. Furthermore, my findings suggest that symbols decrease citizens' "rational", systematic processing of substantive information. This is good news as far as we are concerned mainly about citizens' *undervaluation* of government performance, and the negative biases that undermine the effectiveness of positive information. On the other hand, the emotive effect of government symbols on public opinion is highly problematic, insofar as it challenges the democratic ideal that citizens form their opinions about their government and its actions on the basis of rational thinking [@dahl_1973; @soroka_2010]. Moreover, it becomes especially problematic when symbols are used by poorly performing organizations, or organizations that lack professional expertise. In those cases, symbolic elements can mitigate justified public criticism of government failures. Ultimately, by preserving citizens' trust in them and escaping public blame, organizations can nurture their long-term positive reputation, which provides them with political power and attenuates elected politicians' incentives to monitor and oversight them [@busuioc_2016; @bertelli_2018]. Hence, whereas government communication with citizens is normally perceived as a tool for enhancing democratic responsiveness and accountability, organizations can also use it as a means to undermine them.

@@

##Appendix B: Symbols manipulation
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/symbols_manipulation.png)

@@

*Note: The policy plans communications presented here are the strong policies, in the version adjusted for mobile interface.*

>>>>

---
title: "Can government public communications elicit undue trust? Exploring the interaction between symbols and substantive information in communications"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: apa.csl
urlcolor: blue
---

@@

```{r , include=FALSE, echo=FALSE}
#load("SVIVA_R_ENV.RData")
source("C:/SAAR/UNIVERSITY/R/SVIVA/code/experiment 2/SVIVA_exp2_dm_03.R")

@@

Public administration scholars are highly concerned about the implications of citizens’ low trust in government, and their underestimation of public-service performance. ##ADD more about trust.##

One of the main remedies proposed for this is greater investment in direct government communications to citizens. Congruently, over the past generation, economically-developed democratic countries have witnessed an increase in government public communications, specifically through advertising, social media, and e-government platforms [@eshuis_2012; @marland_2017; @mergel_2013; @mickoleit_2014; @canel_2013]. In the US, for example, federal government agencies spend approximately 1.5 Billion dollars annually on contracts related to public relations and advertising activities in traditional and digital media [@unitedstatesgovernmentaccountabilityoffice_2017].

Extant research on the effect of government public communications on citizens attitudes has mostly focused on the *informative* aspects of these communications, whereby government organizations provide citizens with evidence about their performance and details and explanations about their policies and actions [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @porumbescu_2018; @grimmelikhuijsen_2015; @ho_2017]. This theoretical and empirical focus on information and transparency congrues with the normative view of government public communication as a means for accomplishing democratic accountability, facilitating the effective implementation of their missions, and aligning citizens' perceptions of public organizations with their actual performance [@canel_2013; @lee_2012; @lee_2001; @lee_2008; @lee_2011]. According to this view, public communications are expected to increase citizens' trust in bureaucracies mainly insofar as these communications provide citizens with favorable information on the performance of organizations and logically persuasive explanations for their policies.

@@

@@

Methodologically, the EPM is a good candidate for this study due to its significant investment, over the past years, in public communications. In these communications, the EPM used an array of symbolic elements, which became widely recognized by the public. Here, I focused on three prominent symbolic elements used in the EPM's public communications, and displayed below in **Figure 1**. The first symbolic element is the unique brand logo of the EPM (Figure 1a). This logo consists of a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second symbolic element which I focused on, is the color green, which is strongly associated with environmental protection. The EPM logo and the green color are used in almost all of EPM’s visual communications, and they represent key elements of its unique "visual identity" [@vandenbosch_2005]. The third element is celebrity endorsements. Specifically, I selected two high-profile celebrity comedians, *Tal Friedman* and *Ido Rosenblum*, who previously endorsed two salient advertising campaigns of the EPM. The campaign starring the comedian Tal Friedman (Figure 1b) was launched in 2010, and included a series of advertisements in which Friedman was presented with a head covered in green grass, and accordingly the campaign's slogan was: "starting to think green". The campaign starring the comedian Ido Rosenblum (Figure 1c) was launched in 2017, and focused on reducing the use of disposable bags.[^footnote_campaigns] The two campaigns were highly successful in terms of their recognizability and likability, as suggested by their evaluation reports by the Government advertising agency.

@@

<br>
**Figure 1: Symbolic elements of EPM**
![](C:/SAAR/UNIVERSITY/R/SVIVA/papers/paper2_JPART/revision/myfigures/EPM_symbols.png) 
<font size="0.5">

@@

Altogether, for each policy plan, the survey experiment has a factorial design of 3 (symbols: real symbols, fake-symbols, no-symbols) x 2 (information: strong vs. weak policy). The assignment to the different experimental groups is graphically presented in **Figure 2**. The experimental groups are balanced with respect to demographics, and pre-manipulation variables. Throughout the rest of the paper, I will analyze the results at the policy level, in order to simplify the analyses. Accordingly, I transformed the dataset into a multilevel structure of policies nested within subjects.[^footnote_combined_analyses]

@@

<br>
**Figure 2: Experimental design for policy plan communications**
![](C:/SAAR/UNIVERSITY/R/SVIVA/papers/paper2_JPART/revision/myfigures/experiment_flow.png)

@@

In Model 1.2, the effect of the symbols remains intact, and the effect of the weak policy condition is negative and significant (*Coef* = `r mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_1.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]). In Model 1.3, the interaction between weak policy and real symbols is positive and significant, whereas the interaction with the fake symbols is insignificant. The interpretation of this model suggests that the real symbols had a greater effect on trust when the policy plan was weak (*Coef* = `r mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`] and `r mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_1.3.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`] compared with the no symbols and fake symbols, respectively). When added to the strong policy plan, the effect of the real symbols is weaker, and significant only compared with the no symbols condition (*Coef* = `r mod_1.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)`, *p*<.1, one-tailed). Congruently, the symbols decreased the differences in trust between the strong and weak policies by `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())/(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())*100)%>%round(0)`%, compared with the no symbols group (from `r mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`] to `r mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_1.3.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`]). The interaction between symbols and information significantly contributes to the model fit (Chi2 = `r anova(mod_1.3,mod_1.2)%>%tidy()%>%filter(term=="mod_1.3")%>%select(statistic)%>%max()%>%round(2)`, *p*<.05). To facilitate the interpretation of this interaction, **Figure 3** provides a descriptive demonstration of the effects of the real symbols on trust in policy, under weak and strong information quality.

**Figure 3: The effect of symbolic elements on trust under weak vs. strong information**

@@

The interaction between symbols and information can also be tested *within-subjects*, by comparing the averaged delta across the different symbols conditions. Based on my second hypothesis, I expect that the averaged delta will be lower among subjects in the real symbols condition. This comparison is graphically displayed in **Figure 4**. The averaged delta in the real symbols is significantly lower than in the no symbols and fake symbols control groups (differences are `r (x_realsymbol_delta-x_nosymbol_delta)%>%abs()%>%round(2)` [`r t_delta_symbol_nosymbol[["conf.int"]][1]%>%round(2)`,`r t_delta_symbol_nosymbol[["conf.int"]][2]%>%round(2)`] and `r (x_realsymbol_delta-x_fakesymbol_delta)%>%abs()%>%round(2)` [`r t_delta_symbol_fakesymbol[["conf.int"]][1]%>%round(2)`,`r t_delta_symbol_fakesymbol[["conf.int"]][2]%>%round(2)`], *t*= `r t_delta_symbol_nosymbol[["statistic"]][["t"]]%>%round(3)` and `r t_delta_symbol_fakesymbol[["statistic"]][["t"]]%>%round(3)`, respectively). The differences between the two control groups are insignificant. These results further support my findings above that familiar symbols decrease subjects' inclination to differentiate between the strong and weak policy plans.

**Figure 4: The effect of symbolic elements on within-subject delta between strong and weak policies**

@@

In the following section, I test the postulation that the interaction between symbols and information is explained or mediated by a decrease in elaboration. To reiterate, in the survey, I used reaction time (log) and memory score as measures for participants' level of elaboration of the policy plans. Based on my third hypothesis, I expect to find a negative effect of the real symbols on these two measures. Thereafter, my fourth hypothesis entails moderated mediation relations: the decrease in elaboration mediates the effect of symbols on trust under the weak policy, but not under the strong policy.

@@

In all four models, the coefficients of the real symbols are negative and significant, and the coefficients of the fake symbols are insignificant. The real symbols decreased the number of second subjects spent on the pages of the policy plans in the online survey, compared with both control groups.[^footnote_RT_symbols] The effect is significant at the 95% level compared with the no symbols condition (*Coef* = `r mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_2.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`]), yet it is not sufficiently significant compared with the fake symbols. The averaged memory score of subjects who saw the real symbols was significantly lower, compared with both the no logo and the fake logo groups (*Coef* = `r mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_2.3%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`], and `r mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_2.3.r_1%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`], respectively).[^footnote_memory_symbols] There is no significant change to the results when adding the effects of the information manipulation (in Models 2.2, 2.4). The coefficients of the weak policy in these models suggest that those who saw the fictitious, weak policy plans, spent more time on their pages, yet were less likely to state items that were included in them. **Figure 5** displays the descriptive comparison of the elaboration measures across the symbols conditions. The left side shows its effect on reaction time (logged), and the right side shows its effect on the memory-test scores.

@@

**Figure 5: The effect of symbolic elements on elaboration**

@@

The results of the within-subjects mediation analysis similarly support the moderated mediation hypothesis. The negative effect of the real symbols on the delta within subjects (between trust in the strong and weak policies) is significantly mediated by the negative effect of the real symbols on reaction time and the memory score. With respect to the reaction time, the estimated proportional mediation is `r (table_med_RT__nosymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% compared with the no symbols, and `r (table_med_RT__fakesymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% compared with the fake symbols. Regarding the memory score, the estimated proportional mediation is `r (table_med_memory__nosymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`% and `r (table_med_memory__fakesymbol_delta["Prop. Mediated","Estimate"]*100)%>%round(1)`%, respectively.[^footnote_mediation_sensitivity]

@@

Keeping these limitations in mind, this study has several important implications for public administration research. Most importantly, my findings highlight the significant role of branding and symbolic communication in shaping citizens' attitudes about government. Citizens' responses to symbolic elements in communications thus far received relatively little attention in Public administration literature, compared with the abundant research on their responses to information. Yet, this study, and others, demonstrated that symbols can be fairly effective in increasing citizens' trust, and mitigating their negative attitudes [@alonbarkat_2017; @karens_2016; @teodoro_2018].

Furthermore, my findings suggest that symbols decrease citizens' "rational", systematic processing of substantive information. This is good news as far as we are concerned mainly about citizens' *undervaluation* of government performance, and the negative biases that undermine the effectiveness of positive information [@baekgaard_2016; @hvidman_2016; @james_2014; @james_2017; @marvel_2015; @olsen_2017; @james_2017a; @james_2017b; @christensen_2018; @hvidman_2018]. On the other hand, the emotive effect of government symbols on public opinion is highly problematic, insofar as it challenges the democratic ideal that citizens form their opinions about their government and its actions on the basis of rational thinking [@dahl_1973; @soroka_2010]. Moreover, it becomes especially problematic when symbols are used by poorly performing organizations, or organizations that lack professional expertise. In those cases, symbolic elements can mitigate justified public criticism of government failures. The potency to manipulate public opinion via symbols could be used for promoting the narrow political interests of parties and individual politicians controling the bureaucracies, as suggested by case studies of governemnt public relations [@holtzbacha_2013; @sanders_2013; @young_2013]. At the same time, it could also be used for promoting the interests of the bureaucracies themselves, namely nurturing their long-term positive reputation, which provides them with political power and autonomy [@busuioc_2016; @bertelli_2018]. Hence, whereas government communication with citizens is normally perceived as a tool for enhancing democratic responsiveness and accountability, it can also be use it as a means to undermine them.

@@

##Appendix B: Symbols manipulation
![](C:/SAAR/UNIVERSITY/R/SVIVA/papers/paper2_JPART/revision/myfigures/symbols_manipulation.png)

@@

*Note: The policy plans presented here are the strong policies, in the version adjusted for a mobile interface.*
